### 2017-02-13 thesis/results/results2.tex
tmp commit

Results from previous section confirmed that the pathway metagenes from the \citet{Gatza2010a} study were in the correct directions, and that the pathway metagenes were behaving as expected in different data sets.
Now that the directions of the pathway metagenes were established, these metagenes were ready for comparison with the obesity associted metagenes.
However, all of the pathway metagenes were derived from Gatza's data, whereas the majority of the obesity associted metagenes were derived from Creighton's data, which presents a problem of deciding on which data set the transformation matrices should be generated in.

In theory, the metagenes generated from the application of \gls{svd} and the metagenes generated from transformation matrix would be exactly the same if the transformation matrix was derived from the same data set.
For example, an obesity associted metagene generated from Creighton's data with \gls{svd} would have the same values as the metagenes generated in Creighton's data with the transformation matrix (where the transformation matrix was derived from Creighton's data).
Furthermore, if the \gls{svd}-derived metagenes and transformation matrix-derived metagenes were the same (or at least similar) in a different data set, this suggests that there was no difference in the data sets, at least in terms of the expression of the genetic signatures being investigated.
In other words, the transformation matrix can be made in either the original data set or in a different data set, if there is no difference in the \gls{svd}-derived or transformation matrix-derived metagenes.

% TODO: start from here
To decide which data set the transformation matrices for obesity and pathway associated genetic signatures should be generated in, the \gls{svd}- and transformation matrix(TM)-generated metagene scores were compared in all of the data sets.
As in \cref{sec:pathway_associated_genetic_signatures_from_gatza2010a_study}, all data sets were normalised with the \gls{rma} method and metagenes were ranked based on the number of samples in each data set.
Transformation matrices for the obesity associted genetic signatures were made in Creighton's data set and pathway associted genetic signatures were made in Gatza's data set.
Metagenes for all of the obesity and pathway associated genetic signatures were generated in all of the data sets with \gls{svd} and transformation matrices.
The Spearman correlation of the \gls{svd}-generated and TM-generated metagene scores were calculated for each genetic signatures.

\begin{table}[htpb]
	\centering
	\caption{Spearman correlation of the svd vs tm metagenes}
	\label{tab:svd_vs_tm}
	\begin{tabular}{c}
		gtrmatransres (RMA)
		\hline
		FM & Creighton   &     Cris
		akt   -0.5563 -0.5190 -0.5900
		bcat   0.9905  0.9897  0.9977
		e2f1   0.9193  0.9646  0.8438
		egfr  -0.4040 -0.0430 -0.3358
		er     0.9966  0.9978  0.9942
		her2  -0.9794 -0.9553 -0.5817
		ifna   0.9951 -0.9830 -0.9991
		ifng   0.9718 -0.9086  0.9950
		myc    0.9689  0.9878 -0.9852
		p53    0.7923  0.3808  0.9981
		p63    0.8368  0.8319  0.2951
		pi3k   0.9365 -0.9543  0.5989
		pr     0.9845 -0.9511 -0.9887
		ras   -0.7229 -0.9078 -0.9125
		src    0.9548 -0.9575 -0.7173
		stat3  0.6167 -0.1902  0.9159
		tgfb   0.9890 -0.9918 -0.2543
		tnfa   0.4315 -0.6046  0.9365

			   obsrmatransres (RMA)
		   \hline
		   Gatza        FM       Cris
		rawobs     0.9985 0.9872 -0.9161
		crol       0.9982 0.9926  0.9715
		resobs    -0.9987 0.9898 -0.9652
		rescrol    0.9981 0.9927  0.9571
		caobs      0.9985 0.9893 -0.9468
		cacrol     0.9983 0.9937  0.9677
		caresobs  -0.9988 0.9939  0.9865
		carescrol  0.9984 0.9952  0.9642
		crobs     -0.9928 0.9862 -0.9344

@@

>>>>

Results from previous section confirmed that the pathway metagenes from the \citet{Gatza2010a} study were in the correct directions, and that the pathway metagenes were behaving as expected in different data sets.
Now that the directions of the pathway metagenes were established, these metagenes were ready for comparison with the obesity associated metagenes.
However, all of the pathway metagenes were derived from Gatza's data, whereas the majority of the obesity associated metagenes were derived from Creighton's data, which presents a problem of deciding on which data set the transformation matrices should be generated in.

In theory, the metagenes generated from the application of \gls{svd} and the metagenes generated from transformation matrix would be exactly the same if the transformation matrix was derived from the same data set.
For example, an obesity associated metagene generated from Creighton's data with \gls{svd} would have the same values as the metagenes generated in Creighton's data with the transformation matrix (where the transformation matrix was derived from Creighton's data).
Furthermore, if the \gls{svd}-derived metagenes and transformation matrix-derived metagenes were the same (or at least similar) in a different data set, this suggests that there was no difference in the data sets, at least in terms of the expression of the genetic signatures being investigated.
In other words, the transformation matrix can be made in either the original data set or in a different data set, if there is no difference in the \gls{svd}-derived or transformation matrix-derived metagenes.

% TODO: start from here
To decide which data set the transformation matrices for obesity and pathway associated genetic signatures should be generated in, the \gls{svd}- and transformation matrix(TM)-generated metagene scores were compared in all of the data sets.
As in \cref{sec:pathway_associated_genetic_signatures_from_gatza2010a_study}, all data sets were normalised with the \gls{rma} method and metagenes were ranked based on the number of samples in each data set.
Transformation matrices for the obesity associated genetic signatures were made in Creighton's data set and pathway associated genetic signatures were made in Gatza's data set.
Metagenes for all of the obesity and pathway associated genetic signatures were generated in all of the data sets with \gls{svd} and transformation matrices.
The Spearman correlation of the \gls{svd}-generated and TM-generated metagene scores were calculated for each genetic signatures (\cref{tab:svd_vs_tm_path,tab:svd_vs_tm_obs}).

\begin{table}[htpb]
	\centering
	\caption{Spearman correlation of the svd vs tm metagenes (pathway)}
	\label{tab:svd_vs_tm_path}
	\begin{tabular}{lcccc}
		& Gatza & Creighton & Cris & Fuentes-Mattei\\
		\hline
		\rule{0pt}{2.25ex}Akt & 1.000 & 0.5190 & 0.5900 & 0.5563 \\
		\gls{bcat}            & 1.000 & 0.9897 & 0.9977 & 0.9905 \\
		E2F1                  & 1.000 & 0.9646 & 0.8438 & 0.9193 \\
		\gls{egfr}            & 1.000 & 0.0430 & 0.3358 & 0.4040 \\
		\gls{er}              & 1.000 & 0.9978 & 0.9942 & 0.9966 \\
		\gls{her2}            & 1.000 & 0.9553 & 0.5817 & 0.9794 \\
		\gls{ifna}            & 1.000 & 0.9830 & 0.9991 & 0.9951 \\
		\gls{ifny}            & 1.000 & 0.9086 & 0.9950 & 0.9718 \\
		Myc                   & 1.000 & 0.9878 & 0.9852 & 0.9689 \\
		p53                   & 1.000 & 0.3808 & 0.9981 & 0.7923 \\
		p63                   & 1.000 & 0.8319 & 0.2951 & 0.8368 \\
		\gls{pi3k}            & 1.000 & 0.9543 & 0.5989 & 0.9365 \\
		\gls{pr}              & 1.000 & 0.9511 & 0.9887 & 0.9845 \\
		Ras                   & 1.000 & 0.9078 & 0.9125 & 0.7229 \\
		Src                   & 1.000 & 0.9575 & 0.7173 & 0.9548 \\
		\gls{stat3}           & 1.000 & 0.1902 & 0.9159 & 0.6167 \\
		\gls{tgfb}            & 1.000 & 0.9918 & 0.2543 & 0.9890 \\
		\gls{tnfa}            & 1.000 & 0.6046 & 0.9365 & 0.4315 \\

@@

\begin{table}[htpb]
	\centering
	\caption{Spearman correlation of the svd vs tm metagenes (obesity)}
	\label{tab:svd_vs_tm_obs}
	\begin{tabular}{lcccc}
		& Creighton & Gatza  & Fuentes-Mattei & Cris\\
		\hline
		\rule{0pt}{2.25ex} Cr & 1.000     & 0.9985 & 0.9872 & 0.9161 \\
		Res                   & 1.000     & 0.9987 & 0.9898 & 0.9652 \\
		CrOl                  & 1.000     & 0.9982 & 0.9926 & 0.9715 \\
		ResOl                 & 1.000     & 0.9981 & 0.9927 & 0.9571 \\
		Ca                    & 1.000     & 0.9985 & 0.9893 & 0.9468 \\
		CaRes                 & 1.000     & 0.9988 & 0.9939 & 0.9865 \\
		CaOl                  & 1.000     & 0.9983 & 0.9937 & 0.9677 \\
		CaResOl               & 1.000     & 0.9984 & 0.9952 & 0.9642 \\
		Original              & 1.000     & 0.9928 & 0.9862 & 0.9344 \\
	\end{tabular}
\end{table}

In \cref{tab:svd_vs_tm_path,tab:svd_vs_tm_obs}, the obesity and pathway associated genetic signatures had a correlatin of 1 in Creighton's and Gatza's data sets respectively, as the transformation matrices were generated in those data sets.
The correlations of the obesity and pathway metagenes were variable across different data sets, which was as expected since all the data sets were different from one another.
However, what was unexpected was the fact that there were some genetic signatures that were highly correlated in all of the data sets, whereas the metagene correlation of other signatures were highly variable across different data sets.
For example, the \gls{bcat} pathway metagene was highly correlated in all of the data sets (\textgreater{} 0.98), whereas the \gls{stat3} pathway metagene was variable across different data sets, ranging from 0.1902 to 0.9159 (\cref{tab:svd_vs_tm_path,fig:allmetacor_bar}).
The fact that some pathway associated metagenes were consistent across different data sets suggested that some of these genetic signatures were reliable and did not depend on the data set the transformation matrices were derived from.
On the other hand, the genetic signatures that were not consistent across the data sets were likely to be dependent on the data set in which they were derived from, and therefore the transformation matrices for these signatures must be derived from Gatza's data set.

% TODO: start from here
In contrast to the pathway associated genetic signatures, all of the obesity associated genetic signatures showed high correlation of the metagenes across the different data sets.

\begin{figure}[htpb]
	\centering
	\includegraphics[page=2,width=0.45\linewidth]{results2/allmetacor_bar}
	\includegraphics[page=16,width=0.45\linewidth]{results2/allmetacor_bar}
	\caption{Visual comparison of the correlation of the metagenes across different data sets}
	\label{fig:allmetacor_bar}
\end{figure}
